---
title: "P8105_hw2_wz2722"
author: "wentao zhou"
date: "2024-09-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem.1
```{r}
library(tidyverse)
```


```{r}
transit_data <- read.csv("NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
```

```{r}
transit_clean <- transit_data %>%
  select(Line, Station.Name, Station.Latitude, Station.Longitude, Route1:Route11, 
         Entrance.Type, Entry, Vending, ADA)
transit_clean <- transit_clean%>%
  mutate(Entry = ifelse(Entry == "YES", TRUE, FALSE))

head(transit_clean)
```
#Describe the dataset
This dataset contains information about New York City subway stations, including:
1. Route: The subway line where the station is located.
2. Station name: The name of the station.
3. Longitude and latitude: The geographic coordinates of the station.
4. Service route: The subway line that stops at the station.
5. Entry: A logical variable that indicates whether the entry allows input (TRUE or False).
6. Vending Machine: Indicates whether the station has vending machines.
7. Entrance type: Entrance type (such as revolving gates, etc.).
8. ADA compliance: Whether the station complies with ADA.

For data cleaning, I first select the relevant columns, and then convert the entry variables to logical types by using the mute function and ifelse.

Dimension: After cleaning, the dataset has 1868 rows by 19 columns. These data are very neat because each variable forms a column and each observation forms a row.

#Q1.How many distinct stations are there?
```{r}
distinct_stations <- transit_clean %>%
  distinct(Line, Station.Name) %>%
  nrow()

distinct_stations
```
#Q2.How many stations are ADA compliant?
```{r}
ada_compliant <- transit_clean %>%
  filter(ADA == TRUE) %>%
  distinct(Line, Station.Name) %>%
  nrow()
```

#Q3.What proportion of station entrances / exits without vending allow entrance?
```{r}
entry_withoutvending <- transit_clean %>%
  filter(Vending == "NO", Entry == TRUE) %>%
  nrow()
total_withoutvending <- transit_clean %>%
  filter(Vending == "NO") %>%
  nrow()
proportion_withoutvending <-entry_withoutvending/total_withoutvending
proportion_withoutvending
```

#reformat
```{r}
transit_clean <- transit_clean %>%
  mutate(across(Route1:Route11, as.character))

transit_clean_re <- transit_clean %>%
  pivot_longer(cols = Route1:Route11, names_to = "Route_Number", values_to = "Route") %>%
  filter(!is.na(Route))
#distinct stations serve the A train
The_A_train_stations <- transit_clean_re %>%
  filter(Route == "A") %>%
  distinct(Line, Station.Name) %>%
  nrow()
The_A_train_stations


add_compliant_A <- transit_clean_re %>%
  filter(Route == "A",ADA == "TRUE") %>%
  distinct(Line, Station.Name) %>%
  nrow()

add_compliant_A
```

#Problem.2
```{r}
# Read and clean 
library(readxl)
library(dplyr)
file_path <- "202409 Trash Wheel Collection Data.xlsx"
mr_trash_wheel <- read_excel(file_path, sheet = "Mr. Trash Wheel", skip = 1)
colnames(mr_trash_wheel)
gwynnda_trash_wheel <- read_excel(file_path, sheet = "Gwynnda Trash Wheel", skip = 1)
colnames(gwynnda_trash_wheel)

# Renaming columns 
gwynnda_trash_wheel <- gwynnda_trash_wheel %>%
  rename(Dumpster = 1, Month = 2, Year = 3, Date = 4, Weight_tons = 5, 
         Volume_cubic_yards = 6, Plastic_Bottles = 7, Polystyrene = 8, 
         Cigarette_Butts = 9, Plastic_Bags = 10, Wrappers = 11) %>%
  filter(!is.na(Dumpster)) %>%
  mutate(Sports_Balls = 0, Trash_Wheel = "Gwynnda Trash Wheel")%>%
  mutate(Year = as.character(Year))
gwynnda_trash_wheel

```
#Analyze the Data
#a. Professor Trash Wheel
```{r}
prof_trash_wheel <- read_excel(file_path, sheet = "Professor Trash Wheel", skip = 1)
colnames(prof_trash_wheel)

prof_trash_wheel <- prof_trash_wheel %>%
  rename(Dumpster = 1, Month = 2, Year = 3, Date = 4, Weight_tons = 5, 
         Volume_cubic_yards = 6, Plastic_Bottles = 7, Polystyrene = 8, 
         Cigarette_Butts = 9, Glass_Bottles = 10, Plastic_Bags = 11, 
         Wrappers = 12) %>%
  filter(!is.na(Dumpster)) %>%
  mutate(Sports_Balls = 0, Trash_Wheel = "Professor Trash Wheel")%>%
  mutate(Year = as.character(Year))
prof_trash_wheel
```
#b.Gwynnda 
```{r}
# Read and clean Gwynnda Trash Wheel data
gwynnda_trash_wheel <- read_excel(file_path, sheet = "Gwynnda Trash Wheel", skip = 1)
colnames(gwynnda_trash_wheel)

# Renaming columns based on actual column names
gwynnda_trash_wheel <- gwynnda_trash_wheel %>%
  rename(Dumpster = 1, Month = 2, Year = 3, Date = 4, Weight_tons = 5, 
         Volume_cubic_yards = 6, Plastic_Bottles = 7, Polystyrene = 8, 
         Cigarette_Butts = 9, Plastic_Bags = 10, Wrappers = 11) %>%
  filter(!is.na(Dumpster)) %>%
  mutate(Sports_Balls = 0, Trash_Wheel = "Gwynnda Trash Wheel")%>%
  mutate(Year = as.character(Year))
gwynnda_trash_wheel
```
# Combine the three datasets

```{r}
all_trash_wheels <- bind_rows(mr_trash_wheel %>% mutate(wheel = "Mr. Trash Wheel"),
                              prof_trash_wheel,
                              gwynnda_trash_wheel)
 #Total weight of trash collected by Professor Trash Wheel
total_weight_prof_trash <- combined_trash_wheel_data %>%
  filter(Trash_Wheel == "Professor Trash Wheel") %>%
  summarise(total_weight = sum(Weight_tons, na.rm = TRUE)) %>%
  pull(total_weight)

# Total number of cigarette butts collected by Gwynnda in June 2022
gwynnda_june_2022 <- combined_trash_wheel_data %>%
  filter(Trash_Wheel == "Gwynnda Trash Wheel", Month == "June", Year == 2022) %>%
  summarise(total_butts = sum(Cigarette_Butts, na.rm = TRUE)) %>%
  pull(total_butts)

cat("Total weight of trash collected by Professor Trash Wheel: ", total_weight_prof_trash, " tons\n")
cat("Total number of cigarette butts collected by Gwynnda in June 2022: ", gwynnda_june_2022, "\n")
```


# Problem 3
```{r}
# 加载所需的库
bakers <- read_csv("gbb_datasets/bakers.csv")
bakes <- read_csv("gbb_datasets/bakes.csv")
results <- read_csv("gbb_datasets/results.csv", skip = 2)
viewers <- read_csv("gbb_datasets/viewers.csv")


head(bakers)
head(bakes)
head(results)

# clean and rename
results_clean <- results %>%
  rename(Episode = episode, Series = series, Baker_Name = baker, Result = result)

bakes <- bakes %>%
  mutate(Baker = str_replace_all(Baker, '"', ''))%>%
  rename(Baker_Name = Baker)
  
  

bakers <- bakers %>%
  rename(Baker_Name = 'Baker Name')%>%
  mutate(Baker_Name = word(Baker_Name, 1))
missing_bakers <- anti_join(bakes, bakers, by = c('Series', 'Baker_Name'))

# combine
bake_data <- left_join(bakes, bakers, by = c("Baker_Name", "Series"))

# Merge bake_data with results_clean on Episode and Baker_Name
final_data <- left_join(bake_data, results_clean, by = c("Baker_Name", "Series","Episode"))
```
```{r}
star_bakers <- final_data %>%
  filter(Series >= 5 & Series <= 10, Result %in% c("STAR BAKER", "WINNER")) %>%
  select(Series, Episode, Baker_Name,Result)


# arrange
star_bakers_table <- star_bakers %>%
  arrange(Series, Episode)


viewers_tidy <- viewers_tidy %>%
  mutate(Series = as.numeric(str_replace(Series, "Series ", "")))
viewers_tidy
# Calculate average viewership for Season 1 and Season 5
avg_viewership_s1 <- viewers_tidy %>%
  filter(Series == 1) %>%
  summarize(avg_viewership = mean(Viewership, na.rm = TRUE))

avg_viewership_s5 <- viewers_tidy %>%
  filter(Series == 5) %>%
  summarize(avg_viewership = mean(Viewership, na.rm = TRUE))
```












